### 2015-07-06 narrative/enabling_tech.tex
working in some references

At the lowest level is the Lightweight File Systems (LWFS) project~\cite{lwfs}.
As mentioned above, LWFS's goal is to provide the simplest core services for
a storage system and investigate adding functionality by using auxiliary
services or layering an API on top that offers more functionality. For example,
by eliminating consistency checks from the core system, applications already
managing consistency can eliminate the redundant check enhancing performance.

@@

The LWFS philosophy focuses on providing a universal core with additional
semantics and services layered on top or run along side. For this project, the
Sirocco API is much more complex than a standard POSIX call prompting layering
a more user friendly interface on top. Placing a thin, optimized layer as part
of the storage system will offer the kinds of functionality and semantics this
project proposes. From the user perspective, something like the
ADIOS~\cite{adios} API offers an API nearly as simple as POSIX, but with the
ability to change the actual data transport mechanism without changing the
source code. This affords writing to a single output API while shifting from
writing to a POSIX file system, some middleware like
DataSpaces~\cite{dataspaces}, or some other system like Sirocco all without
changing the source code. ADIOS has proven effective for writing to POSIX file
systems, data staging systems~\cite{predata,dataspaces,nessie}, workflow
systems~\cite{flexpath}, and even nothing through the use of a NULL transport.
Part of this project's goals seeks to understand where the dividing line
between the storage layer shared by all users and middleware being used by a
single or small number of users.

@@

In this project we will extend ADIOS~\cite{adios} with new APIs and novel
techniques to coordinate with the storage back-end. The key idea is offering
the ability to negotiate with the storage system to agree on the tradeoff
between data quality and response time. We will test our ideas on the next
generation OLCF, SNL/LANL, and NERSC systems.

We plan to leverage the many inovations we discovered in the Lightweight File
Systems~\cite{lwfs} (LWFS) project and the Ceph~\cite{weil:osdi06} storage
system project.  In the LWFS project, Sandia sought to demonstrate stripping
down a file system to the bare minimum features required for any storage
system. Then, additional functionality and semantics, such as POSIX-style
directories and consistency requirements, were added in using auxiliary
services separate from the stack or by layering providing a new interface to
the storage system. The project's first phase demonstrated that using a core
object storage system with integrated security providing authentication and
authorization services, customized storage interfaces could be developed with
exactly the overhead required for different applications. The second phase of
LWFS, known as Sirocco~\cite{sirocco}, is investigating how to manage storage
devices for efficient and scalable storage performance.  Sirocco has been in
process for three years and has developed a usable storage infrastructure.

>>>>

At the lowest level is the Lightweight File Systems (LWFS) project~\cite{oldfield:lwfs}.
As mentioned above, LWFS's goal is to provide the simplest core services for
a storage system and investigate adding functionality by using auxiliary
services or layering an API on top that offers more functionality. For example,
by eliminating consistency checks from the core system, applications already
managing consistency can eliminate the redundant check enhancing performance.

@@

The LWFS philosophy focuses on providing a universal core with additional
semantics and services layered on top or run along side. For this project, the
Sirocco API is much more complex than a standard POSIX call prompting layering
a more user friendly interface on top. Placing a thin, optimized layer as part
of the storage system will offer the kinds of functionality and semantics this
project proposes. From the user perspective, something like the
ADIOS~\cite{lofstead:adaptible} API offers an API nearly as simple as POSIX, but with the
ability to change the actual data transport mechanism without changing the
source code. This affords writing to a single output API while shifting from
writing to a POSIX file system, some middleware like
DataSpaces~\cite{dataspaces}, or some other system like Sirocco all without
changing the source code. ADIOS has proven effective for writing to POSIX file
systems, data staging systems~\cite{predata,dataspaces,nessie}, workflow
systems~\cite{flexpath}, and even nothing through the use of a NULL transport.
Part of this project's goals seeks to understand where the dividing line
between the storage layer shared by all users and middleware being used by a
single or small number of users.

@@

In this project we will extend ADIOS~\cite{lofstead:adaptible} with new APIs and novel
techniques to coordinate with the storage back-end. The key idea is offering
the ability to negotiate with the storage system to agree on the tradeoff
between data quality and response time. We will test our ideas on the next
generation OLCF, SNL/LANL, and NERSC systems.

We plan to leverage the many inovations we discovered in the Lightweight File
Systems~\cite{oldfield:lwfs} (LWFS) project and the Ceph~\cite{weil:osdi06} storage
system project.  In the LWFS project, Sandia sought to demonstrate stripping
down a file system to the bare minimum features required for any storage
system. Then, additional functionality and semantics, such as POSIX-style
directories and consistency requirements, were added in using auxiliary
services separate from the stack or by layering providing a new interface to
the storage system. The project's first phase demonstrated that using a core
object storage system with integrated security providing authentication and
authorization services, customized storage interfaces could be developed with
exactly the overhead required for different applications. The second phase of
LWFS, known as Sirocco~\cite{sirocco}, is investigating how to manage storage
devices for efficient and scalable storage performance.  Sirocco has been in
process for three years and has developed a usable storage infrastructure.

### 2015-07-06 narrative/naming-service.tex
working in some references

Structurally, the traditional POSIX naming service offeriing a hierarchical
space consisting of directories and files may be maintained for backwards
compatibility. However, this view will not offer the same strict semantics
POSIX defines. We must break these strict semantics to address scalability
problems forced by the serialized access to a single source for creating and
accessing files.  Several efforts~\cite{giga+,pvfs,others} have worked to
reduce this contention by doing things like reducing the serialized scope to a
single directory or subtree or allowing a single process in a collective file
operation talk with the metadata service and distribute a handle to other
participants. While these approaches help, they do not address this key
scalability limitation of serialized access--even if the metadata service is
spread across multiple nodes. Instead, we will offer a short duration to
consistency POSIX view to offer the performance and consistency requirements an
exascale application demands.

@@

We plan to address this scalability problem by continuing with the LWFS and
Sirocco model. LWFS and Sirocco have take an approach similar to the pure
object stores, but with a focus on the HPC setting. They have abandonded a
fully POSIX compliant metadata service as the default model in favor of a
container/object/fork/addrress tuple for identifying data similar to those used
for pure object stores. By having a service that addresses the object
collections that comprise a single thing, such as a variable or timestep
output, just enough metadata is maintained to make the storage system usable
without additional heavy lifting by clients.  LWFS demonstrated a POSIX-style
namespace on the side kept in sync using a transaction process like
D2T~\cite{d2t} showing that this alternative approach can support traditional
POSIX API calls even though the underlying storage system uses a different
model. Because we are not requiring the entire storage space be addressible
from a single tree root, we can offer multiple independent roots using the
Sirocco object storage. We will investigate how to make the naming and
discovery service scalable under the circumstances outlined above.

>>>>

Structurally, the traditional POSIX naming service offeriing a hierarchical
space consisting of directories and files may be maintained for backwards
compatibility. However, this view will not offer the same strict semantics
POSIX defines. We must break these strict semantics to address scalability
problems forced by the serialized access to a single source for creating and
accessing files.  Several efforts~\cite{patil:2007:giga+,carns:pvfs} have worked to
reduce this contention by doing things like reducing the serialized scope to a
single directory or subtree or allowing a single process in a collective file
operation talk with the metadata service and distribute a handle to other
participants. While these approaches help, they do not address this key
scalability limitation of serialized access--even if the metadata service is
spread across multiple nodes. Instead, we will offer a short duration to
consistency POSIX view to offer the performance and consistency requirements an
exascale application demands.

@@

We plan to address this scalability problem by continuing with the LWFS and
Sirocco model. LWFS and Sirocco have take an approach similar to the pure
object stores, but with a focus on the HPC setting. They have abandonded a
fully POSIX compliant metadata service as the default model in favor of a
container/object/fork/addrress tuple for identifying data similar to those used
for pure object stores. By having a service that addresses the object
collections that comprise a single thing, such as a variable or timestep
output, just enough metadata is maintained to make the storage system usable
without additional heavy lifting by clients.  LWFS demonstrated a POSIX-style
namespace on the side kept in sync using a transaction process like
D2T~\cite{lofstead:2012:txn} showing that this alternative approach can support traditional
POSIX API calls even though the underlying storage system uses a different
model. Because we are not requiring the entire storage space be addressible
from a single tree root, we can offer multiple independent roots using the
Sirocco object storage. We will investigate how to make the naming and
discovery service scalable under the circumstances outlined above.

### 2015-07-06 narrative/qos.tex
working in some references

\paragraph{QoS Background:} 
Quality of Service (QoS) refers to the
properties of the performance quality of a particular service, in
this case the storage system.
%Performance quality can be expressed in either relative or absolute
%terms. Examples of relative performance terms are ``fair'',
%``proportional'', or ``priority/class-based'' (``the higher the
%priority the better the service'' or ``1st class is better than
%2nd''). The key advantage of relative performance terms is that
%they are easy to implement. The key disadvantage is the difficulty
%of making any guarantees based on those implementations --
%even relative guarantees are mired by the well-known effect of
%priority inversion~\cite{lampson:cacm80}. Examples of absolute
%performance terms are ``rate-based'', ``soft real-time'', and
%``hard real-time''. The key advantage of absolute performance
%terms are that they enable the implementation of strong
%guarantees: absolute terms enable agreements between application
%tasks and storage that we call \emph{reservations} and that do not
%change meaning depending
%on the storage system's workload. The key challenge of absolute
%performance terms is that their implementation is non-trivial,
%especially in large-scale systems.
One of the key differentiating features this project offers is driving storage
decisions through QoS \emph{reservations}, rather than continuing the current model
of allowing applications to compete freely without restriction, for part or all of the storage
resources. This traditional approach leads to intereference effects
\cite{lofstead:2010:adaptive,liu_hotstorage} that can greatly impact IO performance predictability. 
%Further, with the introduction of additional tiers in the
%storage stack, the performance variability will increase with greater
%competiion for limited high performance resources. 
%% Temporaily commented out the above sentence, because it may not be true for the next DOE systems. 
Unfettered, this competition can yield a lower system and application performance 
than the case where applications attempt to reserve storage resources, thereby enabling the storage system
to reason over its performance in terms of reservations and actual
performance.
If the system uses reservations intelligently, its performance is predictable and
efficient because resources are not wasted on interference.
In the cases where an appication arrives that needs to run more urgently
than any of the other already running applications, the system
can rapidly respond to it by reducing the amount of resources taken 
by less urgent applications, allowing the urgent application to finish,
and later reverting back to the previous committed resource level for the rest. This can
be automatically enforced via user- (or adiminstrator-) specified \textit{policies}
(section \ref{}).

@@

\paragraph{Related Work:} On current HPC systems, absolute IO performance
guarantees are not possible. Previous work \cite{lofstead:2010:adaptive,liu_hotstorage} 
from this project team uses a client-assisted approach to mitigate IO variablity.
Existing efforts on scheduing resources~\cite{io-cop,dorier} attempt to offer
admission control to maximize storage performance. Unfortunately, these efforts
are limited to participating applications and only applications from a single
platform. For this approach to be effective, the storage system must offer an
approach that handles clients from all connected platforms and does not
require modification to use new storage access APIs. By using a priority
tagging system, we will offer defaults suitable for all applications that can
be informed by simple extensions to the job scheduler or more advanced
management through new APIs.

>>>>

\paragraph{QoS Background:} 
Quality of Service (QoS) refers to the
properties of the performance quality of a particular service, in
this case the storage system.
%Performance quality can be expressed in either relative or absolute
%terms. Examples of relative performance terms are ``fair'',
%``proportional'', or ``priority/class-based'' (``the higher the
%priority the better the service'' or ``1st class is better than
%2nd''). The key advantage of relative performance terms is that
%they are easy to implement. The key disadvantage is the difficulty
%of making any guarantees based on those implementations --
%even relative guarantees are mired by the well-known effect of
%priority inversion~\cite{lampson:cacm80}. Examples of absolute
%performance terms are ``rate-based'', ``soft real-time'', and
%``hard real-time''. The key advantage of absolute performance
%terms are that they enable the implementation of strong
%guarantees: absolute terms enable agreements between application
%tasks and storage that we call \emph{reservations} and that do not
%change meaning depending
%on the storage system's workload. The key challenge of absolute
%performance terms is that their implementation is non-trivial,
%especially in large-scale systems.
One of the key differentiating features this project offers is driving storage
decisions through QoS \emph{reservations}, rather than continuing the current model
of allowing applications to compete freely without restriction, for part or all of the storage
resources. This traditional approach leads to intereference effects
\cite{lofstead:2010:io-variability,liu_hotstorage} that can greatly impact IO performance predictability. 
%Further, with the introduction of additional tiers in the
%storage stack, the performance variability will increase with greater
%competiion for limited high performance resources. 
%% Temporaily commented out the above sentence, because it may not be true for the next DOE systems. 
Unfettered, this competition can yield a lower system and application performance 
than the case where applications attempt to reserve storage resources, thereby enabling the storage system
to reason over its performance in terms of reservations and actual
performance.
If the system uses reservations intelligently, its performance is predictable and
efficient because resources are not wasted on interference.
In the cases where an appication arrives that needs to run more urgently
than any of the other already running applications, the system
can rapidly respond to it by reducing the amount of resources taken 
by less urgent applications, allowing the urgent application to finish,
and later reverting back to the previous committed resource level for the rest. This can
be automatically enforced via user- (or adiminstrator-) specified \textit{policies}
(section \ref{}).

@@

\paragraph{Related Work:} On current HPC systems, absolute IO performance
guarantees are not possible. Previous work \cite{lofstead:2010:io-variability,liu_hotstorage} 
from this project team uses a client-assisted approach to mitigate IO variablity.
Existing efforts on scheduing resources~\cite{io-cop,dorier} attempt to offer
admission control to maximize storage performance. Unfortunately, these efforts
are limited to participating applications and only applications from a single
platform. For this approach to be effective, the storage system must offer an
approach that handles clients from all connected platforms and does not
require modification to use new storage access APIs. By using a priority
tagging system, we will offer defaults suitable for all applications that can
be informed by simple extensions to the job scheduler or more advanced
management through new APIs.
